---
title: "Project"
author: "Kalp"
date: "2024-04-10"
output: html_document
---
```{r}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
obesity.data <- read_excel("~/Downloads/Obesity.transformed_dataset.xlsx")
View(obesity.data)
```


```{r}
obesity.data
```

```{r}
# Method 1: Using is.na() function
missing_values = is.na(obesity.data)
# Summarize missing values by column
colSums(missing_values)
```

```{r}
# Assuming your dataset is named obesity_data
summary(obesity.data)
```
 
```{r}

#Count the frequency of each gender category
gender_counts = table(obesity.data$Gender)

#Create a bar plot
barplot(gender_counts, 
        main = "Distribution of Gender", # Title of the plot
        xlab = "Gender",                 # Label for x-axis
        ylab = "Frequency",              # Label for y-axis
        col = "skyblue",                 # Color of bars
        ylim = c(0, max(gender_counts) * 1.1), # Set ylim to leave space above the tallest bar
        border = NA)                     # Remove border around bars

```

```{r}
# Create a histogram for age
hist(obesity.data$Age,
     main = "Distribution of Age",    # Title of the plot
     xlab = "Age",                    # Label for x-axis
     ylab = "Frequency",              # Label for y-axis
     col = "skyblue",                 # Color of bars
     border = "black")                # Color of border around bars

```

```{r}
# Create a histogram for age
hist(obesity.data$Height,
     main = "Distribution of Height",    # Title of the plot
     xlab = "Age",                    # Label for x-axis
     ylab = "Frequency",              # Label for y-axis
     col = "skyblue",                 # Color of bars
     border = "black")                # Color of border around bars

```


```{r}
# Create a histogram for age
hist(obesity.data$Weight,
     main = "Distribution of Weight",    # Title of the plot
     xlab = "Weight",                    # Label for x-axis
     ylab = "Frequency",              # Label for y-axis
     col = "skyblue",                 # Color of bars
     border = "black")                # Color of border around bars

```

```{r}
par(mfrow = c(2, 2)) # Set up a 2x2 grid for plots

# Bar plot for family_history_with_overweight
barplot(table(obesity.data$family_history_with_overweight),
        main = "Family History with Overweight",
        xlab = "Family History",
        ylab = "Frequency",
        col = "green",
        border = "black")

# Box plot for FAVC

barplot(table(obesity.data$FAVC),
        main = "Consumption of High Caloric food",
        xlab = "FAVC",
        ylab = "Frequency",
        col = "blue",
        border = "black")


barplot(table(obesity.data$SMOKE),
        main = "SMOKE",
        xlab = "SMOKE",
        ylab = "Frequency",
        col = "purple",
        border = "black")

barplot(table(obesity.data$CAEC),
        main = "Consumption of food between meals",
        xlab = "CAEC",
        ylab = "Frequency",
        col = "orange",
        border = "black")
```

```{r}

# Bar plot for SCC
SCC =  table(obesity.data$SCC)
barplot(SCC,
        main = "Calories consumption monitoring",
        xlab = "SCC",
        ylab = "Frequency",
        col = "darkgreen",
        border = "black")


# Box plot for CALC

CALC = table(obesity.data$CALC)
barplot(CALC,  ylab = "Frequency", xlab = "CALC",main = "CALC",col = "blue",  border = "black")

```



```{r}
MTRANS = table (obesity.data$MTRANS)
barplot(MTRANS,
        main = "Mode of Transport",
        xlab = "MTRANS",
        ylab = "Frequency",
        col = "purple",
        border = "black")
```


```{r}
barplot(table(obesity.data$NObeyesdad),
        main = "Obesity level",
        xlab = "NObeyesdad",
        ylab = "Frequency",
        col = "orange",
        border = "black",
        width = 5)
```

```{r}
# Create a histogram for FCVC
barplot(table(obesity.data$FCVC),
     main = "Distribution of Consumption of vegetable",    # Title of the plot
     xlab = "FCVC",                    # Label for x-axis
     ylab = "Frequency",              # Label for y-axis
     col = "skyblue",                 # Color of bars
     border = "black")
```

```{r}
# Create a histogram for NCP
barplot(table(obesity.data$NCP),
     main = "number of main meals",    # Title of the plot
     xlab = "NCP",                    # Label for x-axis
     ylab = "Frequency",              # Label for y-axis
     col = "skyblue",                 # Color of bars
     border = "black")
```

```{r}
# Create a histogram for CH2O
barplot(table(obesity.data$CH2O),
     main = "Distribution of consumption of water",    # Title of the plot
     xlab = "CH2O",                    # Label for x-axis
     ylab = "Frequency",              # Label for y-axis
     col = "skyblue",                 # Color of bars
     border = "black")
```
```{r}
# Create a histogram for FAF
barplot(table(obesity.data$FAF),
     main = "Physical activity frequency",    # Title of the plot
     xlab = "FAF",                    # Label for x-axis
     ylab = "Frequency",              # Label for y-axis
     col = "skyblue",                 # Color of bars
     border = "black")
```

```{r}
# Create a histogram for TUE
barplot(table(obesity.data$TUE),
     main = "Time using technology devices",    # Title of the plot
     xlab = "TUE",                    # Label for x-axis
     ylab = "Frequency",              # Label for y-axis
     col = "skyblue",                 # Color of bars
     border = "black")
```

```{r}
library(ggplot2)

# Explore the categorical variables
categorical_vars <- c("Gender", "family_history_with_overweight", "FAVC","CAEC",  "SMOKE", "SCC", "CALC", "MTRANS", "NObeyesdad")
for (var in categorical_vars) {
  print(ggplot(obesity.data, aes(x = !!sym(var))) + geom_bar() + labs(title = var))
}



# Convert categorical variables to factors
obesity.data$Gender <- as.factor(obesity.data$Gender)
obesity.data$family_history_with_overweight <- as.factor(obesity.data$family_history_with_overweight)
obesity.data$FAVC <- as.factor(obesity.data$FAVC)
obesity.data$CAEC <- as.factor(obesity.data$CAEC)
obesity.data$SMOKE <- as.factor(obesity.data$SMOKE)
obesity.data$SCC <- as.factor(obesity.data$SCC)
obesity.data$CALC <- as.factor(obesity.data$CALC)
obesity.data$MTRANS <- as.factor(obesity.data$MTRANS)
obesity.data$NObeyesdad <- as.factor(obesity.data$NObeyesdad)

```



```{r}
# Barplot of obesity levels
ggplot(obesity.data, aes(x = NObeyesdad)) +
  geom_bar() +
  labs(x = "Obesity Level", y = "Count", title = "Distribution of Obesity Levels")

```




```{r}
# Barplots of categorical variables by obesity level
categorical_vars <- c("Gender", "family_history_with_overweight", "FAVC","FCVC", "NCP","CAEC", "SMOKE","CH2O", "SCC", "FAF","TUE", "CALC", "MTRANS")

for (var in categorical_vars) {
  # Calculate proportions for each category within each level of obesity
  prop_table <- table(obesity.data[[var]], obesity.data$NObeyesdad) / rowSums(table(obesity.data[[var]], obesity.data$NObeyesdad))
  
  # Convert the table to a data frame
  prop_df <- as.data.frame(prop_table)
  
  # Rename columns
  names(prop_df) <- c(var, "NObeyesdad", "Proportion")
  
  # Plot proportions
  plot <- ggplot(prop_df, aes(x = prop_df[[var]], y = Proportion, fill = NObeyesdad)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(x = var, y = "Proportion", title = paste("Proportion of", var, "by Obesity Level"))
  
  print(plot)
}

```

```{r}
# Boxplots of numerical variables by obesity level
numeric_vars <- c("Age", "Height", "Weight", "FCVC", "NCP", "CH2O", "FAF", "TUE")
for (var in numeric_vars) {
  plot <- ggplot(obesity.data, aes(x = NObeyesdad, y = obesity.data[[var]])) +
    geom_boxplot() +
    labs(x = "Obesity Level", y = var, title = paste("Boxplot of", var, "by Obesity Level"))
  print(plot)
}
```



```{r}
install.packages("readr")
library(readr)
install.packages("tidyverse")
library(tidyverse)

```

#stacked bar plot for visualization of association between independent and dependent variable.
```{r}
# Load necessary libraries
library(ggplot2)

# List of categorical independent variables
categorical_vars <- c("family_history_with_overweight", "FAVC", "FCVC", "NCP", "CAEC",
                      "SMOKE", "CH2O", "SCC", "FAF", "TUE", "CALC", "MTRANS")

# Create stacked bar charts for each categorical variable
for (var in categorical_vars) {
  # Create a data frame for plotting
  plot_data <- data.frame(table(obesity.data[[var]], obesity.data$NObeyesdad))
  colnames(plot_data) <- c(var, "NObeyesdad", "Count")
  plot_data$Percentage <- with(plot_data, Count / sum(Count) * 100)
  
  # Plot stacked bar chart
  p <- ggplot(plot_data, aes(x = factor(get(var)), y = Percentage, fill = NObeyesdad)) +
    geom_bar(stat = "identity", position = "fill") +
    labs(x = var, y = "Percentage", title = paste("Stacked Bar Chart of", var, "and NObeyesdad")) +
    theme_minimal() +
    scale_fill_manual(values = c("Under_weight" = "purple", "Normal_Weight" = "yellow", "Pre-Obese" = "skyblue", "Obese" = "pink")) +
    theme(legend.position = "bottom") +
    guides(fill = guide_legend(title = "NObeyesdad"))
  
  # Display the plot
  print(p)
}
```



CHI-SQUARE TEST

H0: There is no association between Obesity and family history with overweight.
H1: There is a significant association between Obesity and family history with overweight.
As p-value(2.2e-16) is less than the significant level, we reject null hypothesis and there is significant association between Obesity and family history with overweight.
```{r}
contingency_table1 = table(obesity.data$NObeyesdad, obesity.data$family_history_with_overweight)
contingency_table1
chi_square_test1 = chisq.test(obesity.data$NObeyesdad, obesity.data$family_history_with_overweight)
chi_square_test1
```

```{r}
library(ggplot2)
```



```{r}
df_contingency <- as.data.frame.table(contingency_table1)

ggplot(data = df_contingency, aes(x = Var2, y = Var1, fill = Freq)) +
  geom_tile() +
  scale_fill_gradient(low = "lightblue", high = "purple") +
  labs(x = "family_history_with_overweight", y = "NObeyesdad") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```



H0: Obesity is independent of the frequent consumption of high caloric food.
H1: Obesity is dependent of the frequent consumption of high caloric food.
As p-value(2.2e-16) is less than the significant level, we reject null hypothesis and there is significant association between Obesity and Frequent consumption of high caloric food.
```{r}
contingency_table2 = table(obesity.data$NObeyesdad, obesity.data$FAVC)
contingency_table2
chi_square_test2 = chisq.test(obesity.data$NObeyesdad, obesity.data$FAVC)
chi_square_test2
```

```{r}
df_contingency <- as.data.frame.table(contingency_table2)

ggplot(data = df_contingency, aes(x = Var2, y = Var1, fill = Freq)) +
  geom_tile() +
  scale_fill_gradient(low = "lightpink", high = "blue") +
  labs(x = "Frequent consumption of high caloric food", y = "NObeyesdad") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


H0: There is no association between Obesity and Frequency of consumption of vegetables.
H1: There is a significant association between Obesity and Frequency of consumption of vegetables.
As p-value(2.2e-16) is less than the significant level, we reject null hypothesis and there is significant association between Obesity and Frequency of consumption of vegetables.
```{r}
contingency_table3 = table(obesity.data$NObeyesdad, obesity.data$FCVC)
contingency_table3
chi_square_test3 = chisq.test(obesity.data$NObeyesdad, obesity.data$FCVC)
chi_square_test3
```

```{r}
df_contingency <- as.data.frame.table(contingency_table3)

ggplot(data = df_contingency, aes(x = Var2, y = Var1, fill = Freq)) +
  geom_tile() +
  scale_fill_gradient(low = "lightblue", high = "purple") +
  labs(x = "Frequency of consumption of vegetables", y = "NObeyesdad") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


H0: There is no association between Obesity and Number of main meals.
H1: There is a significant association between Obesity and Number of main meals.
As p-value(2.2e-16) is less than the significant level, we reject null hypothesis and there is significant association between Obesity and Number of main meals.
```{r}

contingency_table4 = table(obesity.data$NObeyesdad, obesity.data$NCP)
contingency_table4
chi_square_test4 = chisq.test(obesity.data$NObeyesdad, obesity.data$NCP)
chi_square_test4
```

```{r}
df_contingency <- as.data.frame.table(contingency_table4)

ggplot(data = df_contingency, aes(x = Var2, y = Var1, fill = Freq)) +
  geom_tile() +
  scale_fill_gradient(low = "lightblue", high = "purple") +
  labs(x = "Number of main meals", y = "NObeyesdad") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


H0: There is no association between Obesity and Consumption of food between meals.
H1: There is a significant association between Obesity and Consumption of food between meals.
As p-value(2.2e-16) is less than the significant level, we reject null hypothesis and there is significant association between Obesity and Consumption of food between meals.
```{r}
contingency_table5 = table(obesity.data$NObeyesdad, obesity.data$CAEC)
contingency_table5
chi_square_test5 = chisq.test(obesity.data$NObeyesdad, obesity.data$CAEC)
chi_square_test5
```


```{r}
df_contingency <- as.data.frame.table(contingency_table5)

ggplot(data = df_contingency, aes(x = Var2, y = Var1, fill = Freq)) +
  geom_tile() +
  scale_fill_gradient(low = "lightyellow", high = "purple") +
  labs(x = "Consumption of food between meals", y = "NObeyesdad") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


H0: There is no association between Obesity and smoking.
H1: There is a significant association between Obesity and smoking.

```{r}
contingency_table6 = table(obesity.data$NObeyesdad, obesity.data$SMOKE)
contingency_table6
chi_square_test6 = chisq.test(obesity.data$NObeyesdad, obesity.data$SMOKE)
chi_square_test6
```

```{r}
df_contingency <- as.data.frame.table(contingency_table6)


colnames(df_contingency) <- c("Obesity", "Smoke", "Frequency")

# Create a stacked bar plot
ggplot(df_contingency, aes(x = Obesity, y = Frequency, fill = Smoke)) +
  geom_bar(stat = "identity") +
  labs(x = "Obesity Level", y = "Frequency", fill = "Smoking") +
  theme_minimal() +
  ggtitle("Association between Obesity and Smoking")
```

```{r}
df_contingency <- as.data.frame.table(contingency_table6)

ggplot(data = df_contingency, aes(x = Var2, y = Var1, fill = Freq)) +
  geom_tile() +
  scale_fill_gradient(low = "ivory", high = "pink") +
  labs(x = "Smoking", y = "NObeyesdad") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```







H0: There is no association between Obesity and Consumption of water daily.
H1: There is a significant association between Obesity and Consumption of water daily.

```{r}
contingency_table7 = table(obesity.data$NObeyesdad, obesity.data$CH2O)
contingency_table7
chi_square_test7 = chisq.test(obesity.data$NObeyesdad, obesity.data$CH2O)
chi_square_test7
```


```{r}
df_contingency <- as.data.frame.table(contingency_table7)

ggplot(data = df_contingency, aes(x = Var2, y = Var1, fill = Freq)) +
  geom_tile() +
  scale_fill_gradient(low = "lightyellow", high = "purple") +
  labs(x = "Consumption of water daily", y = "NObeyesdad") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


H0: There is no association between Obesity and Calories consumption monitoring.
H1: There is a significant association between Obesity and Calories consumption monitoring.

```{r}
contingency_table8 = table(obesity.data$NObeyesdad, obesity.data$SCC)
contingency_table8
chi_square_test8 = chisq.test(obesity.data$NObeyesdad, obesity.data$SCC)
chi_square_test8
```

```{r}
df_contingency <- as.data.frame.table(contingency_table8)

ggplot(data = df_contingency, aes(x = Var2, y = Var1, fill = Freq)) +
  geom_tile() +
  scale_fill_gradient(low = "lightyellow", high = "purple") +
  labs(x = "Calories consumption monitoring", y = "NObeyesdad") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


H0: There is no association between Obesity and Physical activity frequency.
H1: There is a significant association between Obesity and Physical activity frequency.
As p-value(2.2e-16) is less than the significant level, we reject null hypothesis and there is significant association between Obesity and Physical activity frequency.
```{r}
contingency_table9 = table(obesity.data$NObeyesdad, obesity.data$FAF)
contingency_table9
chi_square_test9 = chisq.test(obesity.data$NObeyesdad, obesity.data$FAF)
chi_square_test9
```


```{r}
df_contingency <- as.data.frame.table(contingency_table9)

ggplot(data = df_contingency, aes(x = Var2, y = Var1, fill = Freq)) +
  geom_tile() +
  scale_fill_gradient(low = "lightgreen", high = "purple") +
  labs(x = "Physical Activity frequency", y = "NObeyesdad") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


H0: There is no association between Obesity and Time using technology devices .
H1: There is a significant association between Obesity and Time using technology devices .

```{r}
contingency_table10 = table(obesity.data$NObeyesdad, obesity.data$TUE)
contingency_table10
chi_square_test10 = chisq.test(obesity.data$NObeyesdad, obesity.data$TUE)
chi_square_test10
```


```{r}
df_contingency <- as.data.frame.table(contingency_table10)

ggplot(data = df_contingency, aes(x = Var2, y = Var1, fill = Freq)) +
  geom_tile() +
  scale_fill_gradient(low = "ivory", high = "steelblue") +
  labs(x = "Time using technology devices", y = "NObeyesdad") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


H0: There is no association between Obesity and Consumption of alcohol.
H1: There is a significant association between Obesity and Consumption of alcohol.

```{r}
contingency_table11 <- table(obesity.data$NObeyesdad, obesity.data$CALC)


fisher_test11 <- fisher.test(contingency_table11, simulate.p.value = TRUE)

print(contingency_table11)

print(fisher_test11)

```



```{r}
df_contingency <- as.data.frame.table(contingency_table11)

ggplot(data = df_contingency, aes(x = Var2, y = Var1, fill = Freq)) +
  geom_tile() +
  scale_fill_gradient(low = "lightblue", high = "purple") +
  labs(x = "Consumption of alcohol", y = "NObeyesdad") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


H0: There is no association between Obesity and Transportation used.
H1: There is a significant association between Obesity and Transportation used.




```{r}
result <- fisher.test(table(obesity.data$NObeyesdad, obesity.data$MTRANS), simulate.p.value = TRUE)

print(result)

```


```{r}
library(ggplot2)
library(tidyr)

contingency_table <- table(obesity.data$NObeyesdad, obesity.data$MTRANS)

df_contingency <- as.data.frame.matrix(contingency_table)

rownames(df_contingency) <- rownames(contingency_table)
colnames(df_contingency) <- colnames(contingency_table)

df_contingency$NObeyesdad <- rownames(df_contingency)

df_melted <- gather(df_contingency, key = "MTRANS", value = "count", -NObeyesdad)

ggplot(data = df_melted, aes(x = MTRANS, y = NObeyesdad, fill = count)) +
  geom_tile() +
  scale_fill_gradient(low = "ivory", high = "lightblue") +
  labs(x = "Transportation used", y = "NObeyesdad", fill = "Count") +
  theme_minimal() +
  ggtitle("Heatmap of Association between Obesity and Mode of Transportation")

```




#ordinal logistic regression for transformed data set.

```{r}
# Install MASS package if not installed
if (!requireNamespace("MASS", quietly = TRUE)) {
  install.packages("MASS")
}

# Load the MASS package
library(MASS)

# Convert 'NObeyesdad' to a factor
obesity.data$NObeyesdad <- factor(obesity.data$NObeyesdad)

# Fit ordinal logistic regression model
ordinal_model <- polr(NObeyesdad ~ ., data = obesity.data)
summary(ordinal_model)

```









```{r}
library(readr)
my_data <- read_csv('original dataset.csv')
```

#from here we work with original datset

```{r}
library(readr)
my_data <- read_csv("original dataset.csv")
my_data
```

```{r}
library(dplyr)


my_data <- my_data %>%
  mutate(NObeyesdad = ifelse(NObeyesdad %in% c("Overweight_Level_I", "Overweight_Level_II"), "pre_obese", NObeyesdad),
         NObeyesdad = ifelse(NObeyesdad %in% c("Obesity_Type_I", "Obesity_Type_II", "Obesity_Type_III"), "obese", NObeyesdad))
View(my_data)
```



```{r}
install.packages("ggplot2")

```
```{r}
library(ggplot2)

```

```{r}
# Load ggplot2 package
library(ggplot2)

# Create a violin plot with a built-in color palette
ggplot(my_data, aes(x = NObeyesdad, y = FCVC, fill = NObeyesdad)) +
  geom_violin() +
  labs(x = "Obesity Levels", y = "Frequency of Consuming Vegetables", 
       title = "Violin Plot of Frequency of Consuming Vegetables by Obesity Levels") +
  scale_fill_brewer(palette = "Set1")

```






```{r}
# Load ggplot2 package
library(ggplot2)

# Create a violin plot with a built-in color palette
ggplot(my_data, aes(x = NObeyesdad, y = NCP, fill = NObeyesdad)) +
  geom_violin() +
  labs(x = "Obesity Levels", y = "Frequency of consumption of main meals", 
       title = "Violin Plot of Frequency of Consuming Main Meals by Obesity Levels") +
  scale_fill_brewer(palette = "Set1")

```
```{r}
# Load ggplot2 package
library(ggplot2)

# Create a violin plot with a built-in color palette
ggplot(my_data, aes(x = NObeyesdad, y = CH2O, fill = NObeyesdad)) +
  geom_violin() +
  labs(x = "Obesity Levels", y = "Frequency of consumption of water", 
       title = "Violin Plot of Frequency of consumption of water") +
  scale_fill_brewer(palette = "Set1")

```
```{r}
# Load ggplot2 package
library(ggplot2)

# Create a violin plot with a built-in color palette
ggplot(my_data, aes(x = NObeyesdad, y = FAF, fill = NObeyesdad)) +
  geom_violin() +
  labs(x = "Obesity Levels", y = "Frequency of Physical activity", 
       title = "Violin Plot of Frequency of Physical activity") +
  scale_fill_brewer(palette = "Set1")

```

```{r}
# Load ggplot2 package
library(ggplot2)

# Create a violin plot with a built-in color palette
ggplot(my_data, aes(x = NObeyesdad, y = TUE, fill = NObeyesdad)) +
  geom_violin() +
  labs(x = "Obesity Levels", y = "Frequency of technology use", 
       title = "Violin Plot of Frequency of technology use") +
  scale_fill_brewer(palette = "Set1")

```

#convert obesity column to binary


```{r}
# NObeyesdad is the original categorical dependent variable representing obesity classification in our dataset
# Create a new binary variable to indicate obese (1) and non-obese (0)
# Assign 1 to categories representing obese individuals and 0 to all other categories
data$new_binary_obesity <- ifelse(data$NObeyesdad %in% c("Obesity_Type_I", "Obesity_Type_II", "Obesity_Type_III"), 1, 0)

# Replace the original dependent variable with the new binary variable in the dataset
#df$obese_binary <- new_binary_obesity

# Remove the original dependent variable from the dataset if needed
#df <- subset(df, select = -c(NObeyesdad))
View(data)
```


#density plot for numerical varaiables


```{r}
library(ggplot2)

# Create conditional density plots for each continuous variable conditioned on the binary outcome
conditional_plots <- lapply(c("Weight", "FCVC", "NCP", "CH2O", "FAF", "TUE"), function(var) {
  ggplot(my_data, aes_string(x = var, fill = factor(new_binary_obesity))) +
    geom_density(alpha = 0.5) +
    labs(title = paste("Conditional Density Plot for", var),
         x = var,
         y = "Density") +
    theme_minimal()
})

# Plot the conditional density plots
gridExtra::grid.arrange(grobs = conditional_plots, ncol = 2)

```
```{r}
# Create the binary variable based on the NObeyesdad variable
my_data$obese_binary <- ifelse(my_data$NObeyesdad %in% c("pre"), 1, 0)

# Check if the binary variable is correctly created
head(my_data$obese_binary)

# Now, create conditional density plots for each continuous variable conditioned on the binary outcome
conditional_plots <- lapply(c("Weight", "FCVC", "NCP", "CH2O", "FAF", "TUE"), function(var) {
  ggplot(my_data, aes_string(x = var, fill = factor(obese_binary))) +
    geom_density(alpha = 0.5) +
    labs(title = paste("Conditional Density Plot for", var),
         x = var,
         y = "Density") +
    theme_minimal()
})

# Plot the conditional density plots
gridExtra::grid.arrange(grobs = conditional_plots, ncol = 2)

```

#box plots to check association between categorical and binary dependent varaible
```{r}
# Load required libraries
library(ggplot2)

# Create boxplots for categorical variables against the binary outcome variable
# Family history with overweight
ggplot(my_data, aes(x = family_history_with_overweight, y = )) +
  geom_boxplot() +
  labs(x = "Family history with overweight", y = "new_binary_obesity")

# FAVC (Frequency of consumption of high caloric food)
ggplot(my_data, aes(x = FAVC, y = new_binary_obesity)) +
  geom_boxplot() +
  labs(x = "FAVC", y = "Obese Binary")

# CAEC (Consumption of food between meals)
ggplot(my_data, aes(x = CAEC, y = new_binary_obesity)) +
  geom_boxplot() +
  labs(x = "CAEC", y = "Obese Binary")

# SMOKE
ggplot(my_data, aes(x = SMOKE, y = new_binary_obesity)) +
  geom_boxplot() +
  labs(x = "SMOKE", y = "Obese Binary")

# SCC (Calories consumption monitoring)
ggplot(my_data, aes(x = SCC, y = new_binary_obesity)) +
  geom_boxplot() +
  labs(x = "SCC", y = "Obese Binary")

# MTRANS (Transportation used)
ggplot(my_data, aes(x = MTRANS, y = new_binary_obesity)) +
  geom_boxplot() +
  labs(x = "MTRANS", y = "Obese Binary")

```




#logitic regression model


```{r}
library(readr)
data <- read_csv("original dataset.csv")
View(data)
```











```{r}
# NObeyesdad is the original categorical dependent variable representing obesity classification in our dataset
# Create a new binary variable to indicate obese (1) and non-obese (0)
# Assign 1 to categories representing obese individuals and 0 to all other categories
data$new_binary_obesity <- ifelse(data$NObeyesdad %in% c("Obesity_Type_I", "Obesity_Type_II", "Obesity_Type_III"), 1, 0)

# Replace the original dependent variable with the new binary variable in the dataset
#df$obese_binary <- new_binary_obesity

# Remove the original dependent variable from the dataset if needed
#df <- subset(df, select = -c(NObeyesdad))
View(data)
```


```{r}
# Fit logistic regression model
model <- glm(new_binary_obesity ~ Gender + Age + Height + Weight + family_history_with_overweight + FAVC + FCVC + NCP + CAEC + SMOKE + CH2O + SCC + FAF + TUE + CALC + MTRANS,
             data = data, family = binomial(link = "logit"))

# Display summary of the model
summary <- summary(model)

```



```{r}
# Now 'data' is our dataframe with the binary dependent variable 'new_binary_obesity' and predictor variables 

# Fit logistic regression model
model <- glm(new_binary_obesity ~ Gender + Age + Height + Weight + family_history_with_overweight + FAVC + FCVC + NCP + CAEC + SMOKE + CH2O + SCC + FAF + TUE + CALC + MTRANS, data = data, family = binomial)

# Summarize the model
summary(model)

# We can plot diagnostic plots to assess the model fit
plot(model)

```


#odds ratio

```{r}
coefficients <- glm(new_binary_obesity ~ Gender + Age + Height + Weight + family_history_with_overweight + FAVC + FCVC + NCP + CAEC + SMOKE + CH2O + SCC + FAF + TUE + CALC + MTRANS, data = data, family = binomial(link = "logit"))$coef

#standard_errors <- summary$coefficients

# Calculate odds ratios
odds_ratios <- exp(coefficients)


results <- data.frame(
  Variable = names(coefficients),odds_ratios
)

# Print results
print(results)
```

```{r}
# Load necessary libraries if not already loaded
# install.packages("tidyverse") # if you haven't installed it yet
library(tidyverse)

# Assuming your data is stored in a dataframe called df
# and your dependent variable is named new_binary_obesity
# and independent variables are Age, family_history_with_overweight, FAVC, SMOKE, SCC, FAF, MTRANS

# Univariate logistic regression for Age
model_age <- glm(new_binary_obesity ~ Age, data = data, family = "binomial")
summary(model_age)

# Univariate logistic regression for family_history_with_overweight
model_family_history <- glm(new_binary_obesity ~ family_history_with_overweight, data = data, family = "binomial")
summary(model_family_history)

# Univariate logistic regression for FAVC
model_FAVC <- glm(new_binary_obesity ~ FAVC, data = data, family = "binomial")
summary(model_FAVC)

# Univariate logistic regression for SMOKE
model_SMOKE <- glm(new_binary_obesity ~ SMOKE, data = data, family = "binomial")
summary(model_SMOKE)

# Univariate logistic regression for SCC
model_SCC <- glm(new_binary_obesity ~ SCC, data = data, family = "binomial")
summary(model_SCC)

# Univariate logistic regression for FAF
model_FAF <- glm(new_binary_obesity ~ FAF, data = data, family = "binomial")
summary(model_FAF)

# Univariate logistic regression for MTRANS
model_MTRANS <- glm(new_binary_obesity ~ MTRANS, data = data, family = "binomial")
summary(model_MTRANS)

```


#new logistic regression model using the variables that showed association with outcome varaible.

```{r}
model2 <- glm(new_binary_obesity ~   Age + family_history_with_overweight + FAVC + SMOKE + SCC + FAF + MTRANS,
             data = data, family = binomial(link = "logit"))

# Display summary of the model
summary <- summary(model2)
summary
```

#predicitve probability

```{r}
# Assuming your model is named model2

# Predictive probabilities for the training data
predicted_probabilities <- predict(model2, type = "response")

# Adding predicted probabilities to your original dataset
data$predicted_probabilities <- predicted_probabilities

# View the predicted probabilities
head(data$predicted_probabilities)

```




```{r}
coefficients <- glm(new_binary_obesity ~   Age + family_history_with_overweight + FAVC + SMOKE + SCC + FAF + MTRANS,
             data = data, family = binomial(link = "logit"))$coef

#standard_errors <- summary$coefficients

# Calculate odds ratios
odds_ratios <- exp(coefficients)


results <- data.frame(
  Variable = names(coefficients),odds_ratios
)

# Print results
print(results)
```



```{r}
library(ggplot2)

# Sort the results dataframe by odds ratios
results <- results[order(results$odds_ratios, decreasing = TRUE), ]

# Calculate confidence intervals
results$ci_upper <- results$odds_ratios + 1.96 * sqrt((1 / results$odds_ratios) * (1 - 1 / results$odds_ratios))
results$ci_lower <- results$odds_ratios - 1.96 * sqrt((1 / results$odds_ratios) * (1 - 1 / results$odds_ratios))

# Create a forest plot
ggplot(results, aes(x = odds_ratios, y = Variable)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray") +
  geom_point(size = 3, color = "blue") +
  geom_errorbarh(aes(xmin = 1 / ci_upper, xmax = ci_upper), height = 0) +
  scale_x_continuous(trans = "log", breaks = c(0.1, 0.5, 1, 2, 5, 10)) +
  labs(x = "Odds Ratio (log scale)", y = NULL, title = "Forest Plot of Odds Ratios") +
  theme_minimal()


```



